Cosmological Hydrodynamic Simulations of Preferential Accretion in the SMBH of Milky Way Size Galaxies

### 2016-09-30 ms.tex
jk THIS is the submitted version

%[Note to self: Cite one canonical person but also more junior people in the field.]
%\citep[for example; ][at the end of the line]{Bellovary2010,Tremmel2015,Fu2008}
 %\S\ref{sec-intro}

@@

% Kelly cut: In addition to the M$-\sigma$, other relations between th,e SMBH and its host have been well studied, making it clear that, though these objects have many orders of magnitude between their sizes, a complex interplay still exists.

%[CITE FOR THEORY: Springel and Dimatteo; cite when we discuss why m-sigma happens]
%McConnell and Ma have the prettiest M-sigma relation, but there are two camps,: Kormendy and Ferrarese & Merritt who must both be cited.
%Check with Kelly: Which Van den Bosch, Emsellem, and Prija papers she would recommend.

% For thesis: M-sigma plot:
%\begin{figure}
%\centerline{\resizebox{0.75\hsize}{!}{\includegraphics[angle=0]{fancymsigma}}}
%\caption[M-$\sigma$ Relation]{The M-$\sigma$ relatpion which depicts the strong correlation between the mass of a central SMBH and the velocity dispersion of its host's stellar population \citep{Xiao2011}.}
%\label{msigma} 
%\end{figure}
% Cite: Xiao

%The black hole mass-bulge luminosity relation was implied by the work of \cite{Dressler1988} and first illustrated by \cite{Kormendy1993}. A wealth of evidence continues to relate these two characteristics, and follow up examinations using HST observations have determined the current, best estimates yield a M$_{\rm BH}$/M$_{\rm bulge}$ fraction between about 0.0013 and 0.0023 \citep{Merritt2001a,McLure2001,Marconi2003}.  Cut per Jillian sept.13.16

@@

%\textcolor{violet}{
It is clear that galaxy hosts grow through a variety of channels that depend on mass, environment, and interaction history. Therefore, we want to understand how these different galaxy evolutionary paths translate into SMBH fueling mechanisms, and see how they affect the fueling gas flowing into the SMBH itself. \cite{Bellovary2013} compared simulations of three high mass, high redshift galaxies and found that while mergers and smooth accretion both efficiently build up galaxies, no particular dynamical process was more adept at feeding the SMBH. However, with only minor mergers, these galaxies represented relatively quiet merger histories. Using a similar method as \cite{Bellovary2013}, this work examines the SMBH and galaxy fueling mechanisms of a MW-mass galaxy with a rich merger history. MW-type galaxies host SMBHs on the order of 10$^6$ $M_{\odot}$, which are likely the most common type of massive black hole, yet little is known about them or how they may grow \citep{Kormendy2013}. Through this examination, we hope to better understand the coevolution of SMBHs and their hosts in this class of galaxy. 
%}

%\textcolor{violet}{
In this study we analyze the Milky Way-type galaxy, h258, which has a history characterized by major mergers. Since this galaxy is similar to the MW in virial mass, stellar mass, and circular velocity, without a deeper examination, we may not recognize the turbulent history from which it results. We will pinpoint the origins of gas entering the SMBH and halo to look for clues about SMBH fueling within this galaxy. By examining its assembly, and its SMBH's fueling, we can determine the accretion rate and gain further understanding about how SMBHs grow over a range of merger histories in galaxies like our own.
%}

@@

% [NOTE TO SELF: Make/include merger trees. Ask Christina?] Can not do easily with current version parameters (halo switch turned off)
%\textcolor{violet}{
Using the smoothed particle hydrodynamics (SPH) N-body tree code, Charm N-body GrAvity solver \citep[ChaNGa;][]{Menon2015}, we ran an initial dark matter-only, uniform resolution volume of 50 h$^{-1}$ Mpc on a side to identify a MW-mass halo at z = 0 for further examination. This DM-only simulation assumed WMAP 3 parameters \citep{Spergel2007}: $\Omega_m$ = 0.24, $\Omega_{\Lambda}$ = 0.76, $H_0$ = 73 km/s, and $\sigma _8$ = 0.77. Halo h258 was chosen for its Milky Way-mass at z=0 and its active merger history. The halo has a virial mass of $M_{\rm vir} = 7 \times 10^{11} M_{\odot}$ at z = 0 defined relative to a critical density, $\rho_c$, where $\rho / \rho_c$ = 100. Two recent major mergers characterize the h258 halo at z=1.8 and z=1.2. We constructed a ``zoom-in'' high resolution simulation on this galaxy, including gas and star particles, using the volume-renormalization of \cite{Katz1993}, resimulating only a few virial radii from the main halo at the highest resolution from z=150 to z=0.  
%}

%\subsection{High Resolution CHANGA Simulations}
%\textcolor{violet}{
We note that a lower resolution version of h258 was run using the Gasoline code \citep{Wadsley2004}. Our higher resolution h258 run has a spline force softening length of 174 pc and initial gas particle masses of 2.7 $\times$ $10^4 M_{\odot}$. Star particles are created with 30$\%$ of their parent gas particle mass, allowing a maximum initial mass of 8100 $M_{\odot}$. Halo h258 contains about 5 million DM particles inside the virial radius at z=0 and over 14 million DM, star, and gas particles total. The resolution of both force and mass in these simulations is comparable to the ``Eris'' simulation which has one of the highest resolutions for an N-body+SPH cosmological simulation of a Milky Way-mass galaxy so far produced \citep{Guedes2011}.

@@

Simulated galaxies are shown to conform with the observed Tully-Fisher relation \citep{Governato2009}, the size-luminosity relation \citep{Brooks2011}, and the mass-metallicity relation \citep{Brooks2007,Christensen2015}, in addition to having realistic matter distributions and baryon fractions \citep{Governato2009a,Guedes2011}. The parameter and resolution choices described above allow the galaxies to adhere to the stellar-mass-halo-mass relation at z=0 and maintain a realistic period of star formation \citep{Moster2010,Munshi2013,Brooks2007,Maiolino2008}. Given the strict adherence of this simulated galaxy with observations, we are confident that it reasonably represents growth in the galaxy and its SMBH.
%}

@@

The requirement that BH seeds must form from zero metallicity gas particles also causes BH formation to be confined in areas of primordial star formation in the earliest and most massive halos in the simulation. In this technique, BH formation is dependent only on local environment, neglecting any large-scale properties of the host halo. We use the sub-grid prescription for modeling the effects of dynamical friction on SMBH orbits from \cite{Tremmel2015}, which has been shown to produce realistic sinking times for SMBHs.  This prescription, combined with our high resolution to minimize two-body interactions and numerical noise, results in SMBHs that can remain stable at galactic center while also, when appropriate, experiencing realistic perturbations and sinking timescales during and after galaxy interactions and mergers \citep{Bellovary2011}.
%Black holes are not fixed within the center of their host, allowing them to be dynamically affected by mergers and other perturbations within the galaxy. Nevertheless, BHs remain near their host centers by choosing dark matter particle masses in high-resolution regions to be on the same order as gas particle masses, minimizing two-body interactions \citep{Bellovary2011}.

@@

%[Talk to Glenna about Ritchie Thomas forces]

@@

%\textcolor{violet}{
We first identify halos using the Amiga Halo Finder which uses an overdensity criterion for a flat universe \citep{Knebe2001,Knollmann2009,Gill2004} to set the virial radius in the primary halo. We select the primary halo to be the most massive at z=0 in the high resolution region. The central SMBH in the primary halo has a mass of $1.3 \times 10^{7} M_{\odot}$.  
%}

@@

%Just as we found in the same galaxy, h258, of the lower resolution simulation, the high resolution h258 galaxy sees a distinct preference for accreting gas that has been gained through the many mergers in its history. In this version,

%\textcolor{violet}{
The galaxy h258 is characterized by two major mergers; the first occurs at z $\sim$ 1.8 (mass ratio, q $\sim$ 0.8) and the second at z $\sim$ 1.2 (q $\sim$ 1). Despite its merger-rich history, gas accretion smoothly increases the cumulative black hole mass in h258 throughout its evolution as can be seen in Figure \ref{hrh258allmassgas}. The black dashed line in \ref{hrh258allmassgas} indicates the total cumulative BH mass (including both mass from gas and BH mergers), while the black solid line indicates the total accreted gas mass. The blue dot-dashed line represents the gas mass accreted via unshocked gas, while the green solid line and red dashed line show the gas mass accreted through mergers and shocked gas, respectively. It is worthwhile to point out that the largest part of the mass budget at high redshift is not gas at all, but other black holes that have merged with the SMBH seed. This has important implications for gravitational wave astronomy, increasing the event rate for SMBH assembly at high redshifts \citep{Holley-Bockelmann2010}. Aside from this early BH assembly, the largest gain in SMBH mass comes from gas associated with the major mergers between z $\sim$ 2 and z $\sim$ 1. It is clear that unshocked gas makes up the majority entering the galaxy at early times. However, the transition between when smooth, unshocked accretion and mergers dominate are clearly distinguished. While clumpy gas (green) dominates at the earliest time, unshocked gas (blue) overtakes it for a short time before clumpy gas once again dominates by z $\sim$ 1.5.

This low redshift transition to a clumpy gas preference results in the large fraction of clumpy gas seen in the SMBH (Figure \ref{hrh258stackfrac}). Figure \ref{hrh258stackfrac} depicts the fractions of total gas in the galaxy and the SMBH at z=0, again differentiated by gas origin. The gas accreted by the galaxy is half (56 $\%$) comprised of merger gas, with 36 $\%$ of the gas entering through unshocked, smooth accretion. The smallest fraction of the total gas is comprised of shocked gas (8 $\%$ ). Unlike the galaxy, nearly three quarters (74 $\%$) of the gas accreted by the central SMBH was accreted via mergers, while only a quarter (24 $\%$ ) is comprised of unshocked, smoothly accreted gas. Shocked gas makes up the last 2$\%$ of total gas entering the SMBH. \textbf{It is evident then that the SMBH more readily accretes gas gained through mergers.} This result is contrary to \cite{Bellovary2013} which found that the fractions of gas comprising the SMBH and its host were nearly the same. Through our results, it appears that when major mergers are a key part of the galaxy assembly history, these mergers may also drive SMBH growth.
%}

%\textcolor{violet}
To examine why mergers feed the SMBH so effectively, we focus on the most major mergers (q \textgreater 0.8), which occur at  z$\sim$1.8 and z$\sim$1.2 and involve the third and second largest halos in the zoom-in simulation, respectively. We also examine the major mergers' solitary effects on the SMBH mass budget. By tracing the gas particles that enter the main halo from these two specific halos, we can determine what mass fraction of those particles would eventually accrete onto the SMBH. We chose a timestep for each merging halo prior to the start of its individual merger and selected the gas particles associated with them. We selected the gas particles associated with the second largest halo at z$\sim$1.5 and the gas particles associated with the third largest halo at z$\sim$2.5. Tracing the gas particles from these two timesteps into their mergers with the main halo, we found that 28$\%$ of total mass budget of the SMBH was comprised of the gas from these major mergers. Considering that 74 $\%$ of the SMBH's accreted gas mass is composed of clumpy gas, this implies that more than a third of the total clumpy gas in the SMBH comes from these two major mergers. From the previous study \citep{Bellovary2013}, we might have thought the SMBH's mass budget would have been comprised of ~56$\%$ clumpy gas since the halo's accreted mass budget is 56$\%$; however, the 28$\%$ of accreted gas entering the SMBH from the secondary and tertiary halos can explain the increase in the overall clumpy fraction of the SMBH's mass budget.
%}

%\textcolor{violet}{
A previous study by \cite{Fu2007} supplies direct observational evidence that gas from a merger can be funneled directly into the vicinity of the SMBH (\textless 1 pc). Their study examined a collection of twelve low-redshift quasars, half of which are characterized by luminous extended emission-line regions (EELRs). These EELRs were found to have metallicities below the mass-metallicity correlation of normal galaxies and are thought to have resulted from massive, galactic superwinds accompanying the creation of the powerful radio jets associated with the quasars. The quasars hosting EELRs were also found to have low-metallicity broad-line regions (BLRs) at their centers, while the quasars without EELRs hosted BLRs with metallicities above Z$_{\odot}$. \cite{Fu2007} determined that the presence of these low-metallicity EELRs as well as the similar, low-metallicity BLRs are evidence of a merger with a gas-rich galaxy. Such an interaction may explain both the infusion of low-metallicity gas into the BLRs and the subsequent ejection of that gas by the radio jets, forming the EELRs.  While our study did not measure metallicity or include radio jets, our overall results support the conclusion of a major merger funneling gas into the center of a galaxy. 
%}

%\textcolor{violet}{
%To better understand the apparent preference for merger-accreted gas, we examine the angular momentum of each type of gas at the moment it enters the galaxy. Figure \ref{hrh258angmom} shows that gas entering the SMBH (dashed lines) has an overall lower angular momentum than gas entering galaxy (solid lines), as expected. This statement is true regardless of the gas state, where green, blue, and red indicate clumpy, unshocked, and shocked gas, respectively. 
%}

%\textcolor{violet}{
%Figure \ref{hrh258angmom_merger1} explicitly shows that during the z~1.8 merger, the lowest angular momentum gas entering the galaxy (solid lines) is merger-driven (green, solid line). This is reflected in the SMBH (green, dashed line) as nearly all the clumpy gas at this time funnels directly into the SMBH. 
%}

@@

%\textcolor{violet}{
To better understand the apparent preference for merger-accreted gas, we examine the angular momentum of the gas at the moment it enters the halo. Figure \ref{hrh258angmom} shows a cumulative distribution of the angular momentum of the gas as it enters the halo (solid lines). We further distinguish the gas that enters the SMBH (dashed lines), still considering its angular at the moment of halo entry. The gas is again distinguished by origin (clumpy, unshocked, shocked being green, blue, and red, respectively). We find that the angular momentum of gas entering which eventually reaches the SMBH is lower overall, and that the lowest angular momentum gas is comprised of both clumpy and unshocked gas. This result can be seen in Figure \ref{hrh258angmom_merger2}, which shows the cumulative distribution of the angular momentum of the incoming gas particles at the time of the most massive merger at z $\sim$ 1.2. (Colors and linestyles as in Figure \ref{hrh258angmom}.) Figure \ref{hrh258angmom_merger2} explicitly shows that the gas ending up in the SMBH enters with the lowest angular momentum. This influx of clumpy gas primarily had its origins in the secondary halo. Examining the angular momentum of all the gas entering from the secondary halo, we found that it was consistent with the angular momentum of the total clumpy gas entering the main halo.  Thus it appears that gas from the secondary galaxy may be directly channeled into the SMBH of the primary, in addition to the standard picture of gas in the host galaxy which loses angular momentum due to torques from the merger dynamics \citep{Capelo2015},

%This may imply that the low angular momentum of the gas accreted onto the SMBH may have some of its angular momentum due to the physical properties of the merger \citep{Capelo2015}.

%}

%\textcolor{violet}{
%We can also examine another similar cumulative distribution which includes specifically angular momentum of the gas that is associated with secondary halo (the second most massive halo in the zoom-in simulation). Figure \ref{hrh258angmom_mergergas} has the same angular momentum information for the main halo (solid lines), but the green dashed line now indicates the angular momentum gas \textit{only} associated with the merging secondary halo. We note that a large portion of this gas enters with a higher angular momentum than the cumulative average, yet not nearly as high as the angular momentum of cumulative unshocked and shocked gas.  Jillian cut 9/13/16
%}

%\textcolor{violet}{
A lower resolution version of the galaxy h258 was run using the N-body code, Gasoline, like the simulations in \cite{Bellovary2013}. Despite fundamental differences in the hydrodynamic implementation and gas physics included and the absence of a dynamical friction prescription, an analysis of this low resolution h258 results in a SMBH with the same distinct preference for accreting clumpy gas. A second galaxy, h277, was also run using Gasoline; however, this galaxy was characterized by a quiescent merger history (no major mergers) and resulted in a SMBH whose accreted mass fractions mirrored the halo (as seen in \cite{Bellovary2013}). The broad consistency between the low and high resolution simulation of the same galaxy, and the similar results of a quiescent galaxy to the previous study, indicates that the large scale gravitational dynamics could be a main driver of the SMBH fueling in this case. We also stress that that while major mergers may not be the only physical mechanisms by which gas can be funneled into the centers of galaxies (the previous study being a strong example of this), mergers between galaxies clearly play an important role when considering the gas accretion of SMBHs.
%}{}

@@

The angular momentum of the accreted gas as it enters the galaxy sheds some light on the mechanism driving this preferentially accreted clumpy gas. Smoothly accreted gas, which enters the galaxy with a wide range of angular momentum, may become additionally torqued by the already existing disk; however, some studies have found that low angular momentum gas can fuel the central SMBH via filemts of smoothly accreted gas \citep{Dubois2012,DiMatteo2016}. Meanwhile, gas entering through mergers is restricted by the direction of its entry and may allow for a small range of low angular momentum gas to enter the main halo. Our study is the first to find that gas accreted through mergers has the advantage of falling more readily to the center and accreting onto the SMBH. Considering all origins of gas, we see a clear distinction wherein lower angular momentum gas preferentially feeds the SMBH.

@@

%\begin{figure}
%\centerline{
 % \resizebox{0.95\hsize}{!}{\includegraphics[angle=0]{fig1a.pdf}}
  %}
%\centerline{
 % \resizebox{0.95\hsize}{!}{\includegraphics[angle=0]{fig1b.pdf}}
  %}  
  %\centerline{
  %\resizebox{0.95\hsize}{!}{\includegraphics[angle=0]{fig1c.pdf}}
  %}
%\caption[]{Evolution of our flat rotating model (A0) in isolation.
%Top Panel: Stellar density profile at various times. $\gamma = 1.0$ is the reference theoretical profile, shown as a dashed line. Middle Panel: Evolution of intermediate to major (b/a) and minor to major %(c/a) axis ratio as measured at the half mass radius. Bottom Panel: Ratio of rotational velocity to the 3-d velocity dispersion as a function of distance at various times. In isolation, the rotational support in the center decreases over time as the central SMBH increases the velocity dispersion.
%} \label{stab}
%\end{figure}

%\begin{figure}
%\centerline{\resizebox{0.75\hsize}{!}{\includegraphics[angle=0]{h258_faceon_crop}}}
%\caption[]{Face-on Sunrise images of our ChaNGa galaxy, h258, which has
%an active, merger-rich history. NEED TO REPLACE, this is low res h258}
%\label{h258face} 
%\end{figure}

%\centerline{\resizebox{0.75\hsize}{!}{\includegraphics[angle=0]{angmom_cumudist_cumu_plusmerger_4096_only1296_andclumpy}}}
%\caption[]{ Cumulative distribution of angular momentum of the gas particles accreted onto the h258 galaxy at the time of the primary major merger (z $\sim$ 1.2). The solid lines indicate the angular momentum of gas accreted onto the main halo (similar to Figure \ref{hrh258angmom}) and the dashed line indicates the angular momentum of gas entering from the secondary halo. As before, green, blue, and red lines indicate clumpy, unshocked, and shocked gas, respectively.}
%\label{hrh258angmom_mergergas} 
%\end{figure}

%\begin{figure}
%\centerline{\resizebox{0.75\hsize}{!}{\includegraphics[angle=0]{hrh258_am_tsnew_1128}}}
%\caption[]{ Cumulative distribution of angular momentum of the gas particles accreted onto the h258 galaxy at the time of the less massive merger (z $\sim$ 1.8). There are about 900,000 gas particles accreted at this timestep, which gas particles accreted onto the main halo and central black hole are distinguished by solid and dashed lines. The green, blue, and red lines indicate clumpy, unshocked, and shocked gas, respectively.}
%\label{hrh258angmom_merger1} 
%\end{figure}

\bibliography{/Users/the_neekster/Documents/REPOSITORIES/Papers/aparticularappetite2016/Sanchez2016.bib}

>>>>

@@

@@

It is clear that galaxy hosts grow through a variety of channels that depend on mass, environment, and interaction history. Therefore, we want to understand how these different galaxy evolutionary paths translate into SMBH fueling mechanisms, and see how they affect the fueling gas flowing into the SMBH itself. \cite{Bellovary2013} compared simulations of three high mass, high redshift galaxies and found that while mergers and smooth accretion both efficiently build up galaxies, no particular dynamical process was more adept at feeding the SMBH. However, with only minor mergers, these galaxies represented relatively quiet merger histories. Using a similar method as \cite{Bellovary2013}, this work examines the SMBH and galaxy fueling mechanisms of a MW-mass galaxy with a rich merger history. MW-type galaxies host SMBHs on the order of 10$^6$ $M_{\odot}$, which are likely the most common type of massive black hole, yet little is known about them or how they may grow \citep{Kormendy2013}. Through this examination, we hope to better understand the coevolution of SMBHs and their hosts in this class of galaxy.

In this study we analyze the Milky Way-type galaxy, h258, which has a history characterized by major mergers. Since this galaxy is similar to the MW in virial mass, stellar mass, and circular velocity, without a deeper examination, we may not recognize the turbulent history from which it results. We will pinpoint the origins of gas entering the SMBH and halo to look for clues about SMBH fueling within this galaxy. By examining its assembly, and its SMBH's fueling, we can determine the accretion rate and gain further understanding about how SMBHs grow over a range of merger histories in galaxies like our own.

@@

Using the smoothed particle hydrodynamics (SPH) N-body tree code, Charm N-body GrAvity solver \citep[ChaNGa;][]{Menon2015}, we ran an initial dark matter-only, uniform resolution volume of 50 h$^{-1}$ Mpc on a side to identify a MW-mass halo at z = 0 for further examination. This DM-only simulation assumed WMAP 3 parameters \citep{Spergel2007}: $\Omega_m$ = 0.24, $\Omega_{\Lambda}$ = 0.76, $H_0$ = 73 km/s, and $\sigma _8$ = 0.77. Halo h258 was chosen for its Milky Way-mass at z=0 and its active merger history. The halo has a virial mass of $M_{\rm vir} = 7 \times 10^{11} M_{\odot}$ at z = 0 defined relative to a critical density, $\rho_c$, where $\rho / \rho_c$ = 100. Two recent major mergers characterize the h258 halo at z=1.8 and z=1.2. We constructed a ``zoom-in'' high resolution simulation on this galaxy, including gas and star particles, using the volume-renormalization of \cite{Katz1993}, resimulating only a few virial radii from the main halo at the highest resolution from z=150 to z=0.

We note that a lower resolution version of h258 was run using the Gasoline code \citep{Wadsley2004}. Our higher resolution h258 run has a spline force softening length of 174 pc and initial gas particle masses of 2.7 $\times$ $10^4 M_{\odot}$. Star particles are created with 30$\%$ of their parent gas particle mass, allowing a maximum initial mass of 8100 $M_{\odot}$. Halo h258 contains about 5 million DM particles inside the virial radius at z=0 and over 14 million DM, star, and gas particles total. The resolution of both force and mass in these simulations is comparable to the ``Eris'' simulation which has one of the highest resolutions for an N-body+SPH cosmological simulation of a Milky Way-mass galaxy so far produced \citep{Guedes2011}.

@@

Simulated galaxies are shown to conform with the observed Tully-Fisher relation \citep{Governato2009}, the size-luminosity relation \citep{Brooks2011}, and the mass-metallicity relation \citep{Brooks2007,Christensen2015}, in addition to having realistic matter distributions and baryon fractions \citep{Governato2009a,Guedes2011}. The parameter and resolution choices described above allow the galaxies to adhere to the stellar-mass-halo-mass relation at z=0 and maintain a realistic period of star formation \citep{Moster2010,Munshi2013,Brooks2007,Maiolino2008}. Given the strict adherence of this simulated galaxy with observations, we are confident that it reasonably represents growth in the galaxy and its SMBH.

@@

The requirement that BH seeds must form from zero metallicity gas particles also causes BH formation to be confined in areas of primordial star formation in the earliest and most massive halos in the simulation. In this technique, BH formation is dependent only on local environment, neglecting any large-scale properties of the host halo. We use the sub-grid prescription for modeling the effects of dynamical friction on SMBH orbits from \cite{Tremmel2015}, which has been shown to produce realistic sinking times for SMBHs.  This prescription, combined with our high resolution to minimize two-body interactions and numerical noise, results in SMBHs that can remain stable at galactic center while also, when appropriate, experiencing realistic perturbations and sinking timescales during and after galaxy interactions and mergers \citep{Bellovary2011}.

@@

@@

We first identify halos using the Amiga Halo Finder which uses an overdensity criterion for a flat universe \citep{Knebe2001,Knollmann2009,Gill2004} to set the virial radius in the primary halo. We select the primary halo to be the most massive at z=0 in the high resolution region. The central SMBH in the primary halo has a mass of $1.3 \times 10^{7} M_{\odot}$.

@@

The galaxy h258 is characterized by two major mergers; the first occurs at z $\sim$ 1.8 (mass ratio, q $\sim$ 0.8) and the second at z $\sim$ 1.2 (q $\sim$ 1). Despite its merger-rich history, gas accretion smoothly increases the cumulative black hole mass in h258 throughout its evolution as can be seen in Figure \ref{hrh258allmassgas}. The black dashed line in \ref{hrh258allmassgas} indicates the total cumulative BH mass (including both mass from gas and BH mergers), while the black solid line indicates the total accreted gas mass. The blue dot-dashed line represents the gas mass accreted via unshocked gas, while the green solid line and red dashed line show the gas mass accreted through mergers and shocked gas, respectively. It is worthwhile to point out that the largest part of the mass budget at high redshift is not gas at all, but other black holes that have merged with the SMBH seed. This has important implications for gravitational wave astronomy, increasing the event rate for SMBH assembly at high redshifts \citep{Holley-Bockelmann2010}. Aside from this early BH assembly, the largest gain in SMBH mass comes from gas associated with the major mergers between z $\sim$ 2 and z $\sim$ 1. It is clear that unshocked gas makes up the majority entering the galaxy at early times. However, the transition between when smooth, unshocked accretion and mergers dominate are clearly distinguished. While clumpy gas (green) dominates at the earliest time, unshocked gas (blue) overtakes it for a short time before clumpy gas once again dominates by z $\sim$ 1.5.

This low redshift transition to a clumpy gas preference results in the large fraction of clumpy gas seen in the SMBH (Figure \ref{hrh258stackfrac}). Figure \ref{hrh258stackfrac} depicts the fractions of total gas in the galaxy and the SMBH at z=0, again differentiated by gas origin. The gas accreted by the galaxy is half (56 $\%$) comprised of merger gas, with 36 $\%$ of the gas entering through unshocked, smooth accretion. The smallest fraction of the total gas is comprised of shocked gas (8 $\%$ ). Unlike the galaxy, nearly three quarters (74 $\%$) of the gas accreted by the central SMBH was accreted via mergers, while only a quarter (24 $\%$ ) is comprised of unshocked, smoothly accreted gas. Shocked gas makes up the last 2$\%$ of total gas entering the SMBH. \textbf{It is evident then that the SMBH more readily accretes gas gained through mergers.} This result is contrary to \cite{Bellovary2013} which found that the fractions of gas comprising the SMBH and its host were nearly the same. Through our results, it appears that when major mergers are a key part of the galaxy assembly history, these mergers may also drive SMBH growth.

To examine why mergers feed the SMBH so effectively, we focus on the most major mergers (q \textgreater 0.8), which occur at  z$\sim$1.8 and z$\sim$1.2 and involve the third and second largest halos in the zoom-in simulation, respectively. We also examine the major mergers' solitary effects on the SMBH mass budget. By tracing the gas particles that enter the main halo from these two specific halos, we can determine what mass fraction of those particles would eventually accrete onto the SMBH. We chose a timestep for each merging halo prior to the start of its individual merger and selected the gas particles associated with them. We selected the gas particles associated with the second largest halo at z$\sim$1.5 and the gas particles associated with the third largest halo at z$\sim$2.5. Tracing the gas particles from these two timesteps into their mergers with the main halo, we found that 28$\%$ of total mass budget of the SMBH was comprised of the gas from these major mergers. Considering that 74 $\%$ of the SMBH's accreted gas mass is composed of clumpy gas, this implies that more than a third of the total clumpy gas in the SMBH comes from these two major mergers. From the previous study \citep{Bellovary2013}, we might have thought the SMBH's mass budget would have been comprised of ~56$\%$ clumpy gas since the halo's accreted mass budget is 56$\%$; however, the 28$\%$ of accreted gas entering the SMBH from the secondary and tertiary halos can explain the increase in the overall clumpy fraction of the SMBH's mass budget.

A previous study by \cite{Fu2007} supplies direct observational evidence that gas from a merger can be funneled directly into the vicinity of the SMBH (\textless 1 pc). Their study examined a collection of twelve low-redshift quasars, half of which are characterized by luminous extended emission-line regions (EELRs). These EELRs were found to have metallicities below the mass-metallicity correlation of normal galaxies and are thought to have resulted from massive, galactic superwinds accompanying the creation of the powerful radio jets associated with the quasars. The quasars hosting EELRs were also found to have low-metallicity broad-line regions (BLRs) at their centers, while the quasars without EELRs hosted BLRs with metallicities above Z$_{\odot}$. \cite{Fu2007} determined that the presence of these low-metallicity EELRs as well as the similar, low-metallicity BLRs are evidence of a merger with a gas-rich galaxy. Such an interaction may explain both the infusion of low-metallicity gas into the BLRs and the subsequent ejection of that gas by the radio jets, forming the EELRs.  While our study did not measure metallicity or include radio jets, our overall results support the conclusion of a major merger funneling gas into the center of a galaxy.

@@

To better understand the apparent preference for merger-accreted gas, we examine the angular momentum of the gas at the moment it enters the halo. Figure \ref{hrh258angmom} shows a cumulative distribution of the angular momentum of the gas as it enters the halo (solid lines). We further distinguish the gas that enters the SMBH (dashed lines), still considering its angular at the moment of halo entry. The gas is again distinguished by origin (clumpy, unshocked, shocked being green, blue, and red, respectively). We find that the angular momentum of gas entering which eventually reaches the SMBH is lower overall, and that the lowest angular momentum gas is comprised of both clumpy and unshocked gas. This result can be seen in Figure \ref{hrh258angmom_merger2}, which shows the cumulative distribution of the angular momentum of the incoming gas particles at the time of the most massive merger at z $\sim$ 1.2. (Colors and linestyles as in Figure \ref{hrh258angmom}.) Figure \ref{hrh258angmom_merger2} explicitly shows that the gas ending up in the SMBH enters with the lowest angular momentum. This influx of clumpy gas primarily had its origins in the secondary halo. Examining the angular momentum of all the gas entering from the secondary halo, we found that it was consistent with the angular momentum of the total clumpy gas entering the main halo.  Thus it appears that gas from the secondary galaxy may be directly channeled into the SMBH of the primary, in addition to the standard picture of gas in the host galaxy which loses angular momentum due to torques from the merger dynamics \citep{Capelo2015},

A lower resolution version of the galaxy h258 was run using the N-body code, Gasoline, like the simulations in \cite{Bellovary2013}. Despite fundamental differences in the hydrodynamic implementation and gas physics included and the absence of a dynamical friction prescription, an analysis of this low resolution h258 results in a SMBH with the same distinct preference for accreting clumpy gas. A second galaxy, h277, was also run using Gasoline; however, this galaxy was characterized by a quiescent merger history (no major mergers) and resulted in a SMBH whose accreted mass fractions mirrored the halo (as seen in \cite{Bellovary2013}). The broad consistency between the low and high resolution simulation of the same galaxy, and the similar results of a quiescent galaxy to the previous study, indicates that the large scale gravitational dynamics could be a main driver of the SMBH fueling in this case. We also stress that that while major mergers may not be the only physical mechanisms by which gas can be funneled into the centers of galaxies (the previous study being a strong example of this), mergers between galaxies clearly play an important role when considering the gas accretion of SMBHs.

@@

The angular momentum of the accreted gas as it enters the galaxy sheds some light on the mechanism driving this preferentially accreted clumpy gas. Smoothly accreted gas, which enters the galaxy with a wide range of angular momentum, may become additionally torqued by the already existing disk; however, some studies have found that low angular momentum gas can fuel the central SMBH via filemts of smoothly accreted gas \citep{Dubois2012,DiMatteo2016}. Meanwhile, gas entering through mergers is restricted by the direction of its entry and may allow for a small range of low angular momentum gas to enter the main halo. This restriction gives gas accreted through mergers the advantage of falling more readily to the center and accreting onto the SMBH. Considering all origins of gas, our study is the first to see a clear distinction wherein lower angular momentum gas from major mergers preferentially feeds the SMBH.

@@

\bibliography{Sanchez2016.bib}
